---
title: "gtsummary in quarto"
format: html
execute:
  cache: refresh
---

```{r}
#| echo: false
#| output: false
## Alt option to skip chunk entirely would be include: false

library(tidyverse)
library(gtsummary)

nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd", 
               "id", "nsibs", "samp", "race_eth", "sex", "region", 
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"), 
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |> 
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes"))) 
```

# Customization of `tbl_summary()`

```{r}
#| error: true
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

#### EXERCISE 2.3: Quarto tables, figures and stats

1.  Choose a table to label and caption, then write a sentence that cross-references it.
2.  From that table, choose at least 2 statistics to pull out of the table and include in the text using inline_text()
3.  Add another statistic to the text that you calculate yourself using the nlsy data
4.  Render the document, make sure it renders and then commit & push.

```{r}
#| label: tbl-one
#| tbl-cap: "**Table 1.** Descriptive statistics of National Longitudinal Survey of Youth sample population."

table1 <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing") |> 
  add_p(test = list(all_continuous() ~ "t.test", 
                    all_categorical() ~ "chisq.test")) |> 
  add_overall(col_label = "**Total**") |> 
  bold_labels() |> 
  modify_footnote(update = everything() ~ NA) |> 
  modify_header(label = "**Variable**", p.value = "**P**")

table1

```

@tbl-one shows descriptive statistics for the study sample.

*Option 1. Reference results using inline_text function within sentence itself. (Harder to read)*

A significantly greater proportion of females (`r inline_text(table1, variable = "glasses", column = "Female", pattern = "{p}%")`) than males (`r inline_text(table1, variable = "glasses", column = "Male", pattern = "{p}%")`) reported wearing glasses.

*Option 2. Store stats in a list before referencing in text to make it easier to read the inline references*
```{r}
#| output: false
glasses_f <- inline_text(table1, variable = "glasses", column = "Female", pattern = "{p}%")
glasses_m <- inline_text(table1, variable = "glasses", column = "Male", pattern = "{p}%")
```

A significantly greater proportion of females (`r glasses_f`) than males (`r glasses_m`) reported wearing glasses. 


```{r}
##NEED TO DO #3 OF THIS EXERCISE
```



Tips for formatting the number of digits in your inline reference:
- use the round() function to specify the number of significant digits when saving the output to a list
- use the round() function in your inline reference to modify how the number saved in the list is displayed
- Note: round() function can have formating issues if the last digit is a zero. For example, if you want 2 decimal places and the answer is 6.60 it will only show 6.6.
- To avoid this formating challenge, Lousia recommends the scales package and the function number() 
- Link to scales package: https://scales.r-lib.org/reference/number_format.html



# Univariate regression

```{r}
tbl_uvregression(
  nlsy, 
  y = income,
  include = c(sex_cat, race_eth_cat,
              eyesight_cat, income, age_bir),
  method = lm)
```

```{r}
tbl_uvregression(
  nlsy, 
  y = glasses,
  include = c(sex_cat, race_eth_cat,
              eyesight_cat, glasses, age_bir),
  method = glm,
  method.args = list(family = binomial()),
  exponentiate = TRUE)
```

## Multivariable regressions

## Some regressions

```{r}
linear_model <- lm(income ~ sex_cat + age_bir + race_eth_cat, 
                   data = nlsy)
```

```{r}
linear_model_int <- lm(income ~ sex_cat*age_bir + race_eth_cat, 
                   data = nlsy)
```

```{r}
logistic_model <- glm(glasses ~ eyesight_cat + sex_cat + income, 
                      data = nlsy, family = binomial())
```

## Tables

```{r}
tbl_regression(
  linear_model, 
  intercept = TRUE,
  label = list(
    sex_cat ~ "Sex",
    race_eth_cat ~ "Race/ethnicity",
    age_bir ~ "Age at first birth"
  ))
```

```{r}
tbl_regression(
  logistic_model, 
  exponentiate = TRUE,
  label = list(
    sex_cat ~ "Sex",
    eyesight_cat ~ "Eyesight",
    income ~ "Income"
  ))
```

```{r}
tbl_no_int <- tbl_regression(
  linear_model, 
  intercept = TRUE,
  label = list(
    sex_cat ~ "Sex",
    race_eth_cat ~ "Race/ethnicity",
    age_bir ~ "Age at first birth"
  ))

tbl_int <- tbl_regression(
  linear_model_int, 
  intercept = TRUE,
  label = list(
    sex_cat ~ "Sex",
    race_eth_cat ~ "Race/ethnicity",
    age_bir ~ "Age at first birth",
    `sex_cat:age_bir` ~ "Sex/age interaction"
  ))
```

```{r}
tbl_merge(list(tbl_no_int, tbl_int), 
          tab_spanner = c("**Model 1**", "**Model 2**"))
```
